Pollution, agricultural productivity, and development: Evidence from coal in plants in India\footnote{All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}

### 2023-04-12 draft/draft.Rmd
updating some figures

---
title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
author:
  - Joshua D. Merfeld^[KDI School of Public Policy and Management and IZA; merfeld@kdis.ac.kr]
date: "`r Sys.Date()`"
abstract: |
  | \noindent I document large negative externalities in the agricultural sector of India due to air pollution. Using random variation in pollution exposure induced by changes in wind direction, I show that higher levels of particulate matter lead to decreased agricultural productivity. Based on within-village average deviations in pollution exposure, year-to-year changes in productivity can be above 10 percent. Combining wind direction with the rollout of coal plants, I show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution. Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.
  |
  | \textbf{\textit{Keywords}}: pollution, productivity, agriculture, labor, India
  | \textbf{\textit{JEL Codes}}: H40, I15, J22, O13, Q52, Q53

@@

library(sf)
library(sp)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(ggthemes)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(viridis)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

@@

The rest of the paper is organized as follows. I first discuss data and methodology. I break this section up into \autoref{data2}, which covers the multiple sources of data used in this paper; \autoref{identification}, which discusses the empirical strategy and validates the wind direction variable using data on particulate matter. I present the main results in \autoref{results} before concluding in \autoref{conclusion}.

@@

p1 <- ggplot() + 
        theme_map() +
        geom_sf(data = india_shape, fill = NA) +
        geom_tile(data = windexample, aes(x = x, y = y, fill = wind_example), alpha = 0.3) +
        scale_fill_viridis_c("wind direction\n(degrees)", option = "viridis") +
        geom_sf(data = plants2010, size = 0.5) +
        labs(x = "lon",
             y = "lat",
             subtitle = "A. Wind direction") +
        theme(legend.position = c(0.6, 0.05),
              legend.direction = "vertical",
              legend.key.height = unit(0.4, 'cm'),
              legend.text = element_text(size = 8),
              legend.background=element_blank() )

p2 <- ggplot() + 
        theme_map() +
        geom_sf(data = districts, fill = NA) +
        geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
        scale_fill_viridis_c("exposure (days)", option = "viridis") +
        labs(x = "lon",
             y = "lat",
             subtitle = "B. Aggregation to district") +
        theme(legend.position = c(0.5, 0.05),
              legend.direction = "horizontal",
              legend.key.width = unit(0.3, 'cm'),
              legend.text = element_text(size = 8),
              legend.background=element_blank() )

>>>>

---
title: "Air pollution and agricultural productivity in a developing country^[I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution_development]"
author:
  - Joshua D. Merfeld^[KDI School of Public Policy and Management and IZA; merfeld@kdis.ac.kr]
date: "`r Sys.Date()`"
abstract: |
  | \noindent I document externalities in the agricultural sector of India due to air pollution. Using variation in pollution induced by changes in wind, I show that higher levels of particulate matter lead to decreased agricultural productivity, with changes in productivity of 10 percent being common. Combining wind direction with the rollout of coal plants, results indicate pollution from coal plants has a larger effect on agricultural productivity than other types of pollution. Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.
  |
  | \textbf{\textit{Keywords}}: pollution, productivity, agriculture, labor, India
  | \textbf{\textit{JEL Codes}}: H40, I15, J22, O13, Q52, Q53

@@

library(sf)
library(sp)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(ggthemes)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(viridis)
library(stars)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

@@

The rest of the paper is organized as follows. I first discuss data and then methodology, in two separate sections. The methodology section discusses the empirical strategy and validates the wind direction variable using data on particulate matter. I present the main results in \autoref{results} before concluding in \autoref{conclusion}.

@@

windexample2 <- rasterFromXYZ(windexample)
windexample2 <- mask(windexample2, india_shape)
windexample2 <- rasterToPoints(windexample2)
windexample2 <- as.data.frame(windexample2)
p1 <- ggplot() + 
        theme_map() +
        geom_sf(data = india_shape, fill = NA) +
        geom_tile(data = windexample2, aes(x = x, y = y, fill = wind_example), alpha = 0.3) +
        scale_fill_viridis_c("wind direction\n(degrees)", option = "viridis") +
        geom_sf(data = plants2010, size = 0.5) +
        labs(x = "lon",
             y = "lat",
             subtitle = "A. Wind direction") +
        theme(legend.position = c(0.6, 0.05),
              legend.direction = "vertical",
              legend.key.height = unit(0.4, 'cm'),
              legend.text = element_text(size = 8),
              legend.background=element_blank() )

p2 <- ggplot() + 
        theme_map() +
        geom_sf(data = districts, fill = NA) +
        geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
        scale_fill_viridis_c("exposure (days)", option = "viridis") +
        labs(x = "lon",
             y = "lat",
             subtitle = "B. Aggregation to district") +
        theme(legend.position = c(0.45, 0.05),
              legend.direction = "horizontal",
              legend.key.width = unit(0.8, 'cm'),
              legend.text = element_text(size = 8),
              legend.background=element_blank() ) +
        guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5))

### 2023-04-12 draft/draft.tex
updating some figures

\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
\author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
\date{2023-04-11}

\begin{document}
\maketitle
\begin{abstract}
\noindent I document large negative externalities in the agricultural sector of India due to air pollution. Using random variation in pollution exposure induced by changes in wind direction, I show that higher levels of particulate matter lead to decreased agricultural productivity. Based on within-village average deviations in pollution exposure, year-to-year changes in productivity can be above 10 percent. Combining wind direction with the rollout of coal plants, I show that pollution from coal plants, specifically, has a larger effect on agricultural productivity than other types of pollution. Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.\\
\strut \\
\textbf{\textit{Keywords}}: pollution, productivity, agriculture, labor, India\\
\textbf{\textit{JEL Codes}}: H40, I15, J22, O13, Q52, Q53
\end{abstract}

>>>>

\title{Air pollution and agricultural productivity in a developing country\footnote{I thank Ariel Ortiz-Bobea, Marc Bellemare, and David Sungho Park for comments and suggestions that greatly improved the quality of the paper. The usual disclaimer applies. All scripts used in this paper are publicly available on my GitHub page: github.com/JoshMerfeld/pollution\_development}}
\author{Joshua D. Merfeld\footnote{KDI School of Public Policy and Management and IZA; \href{mailto:merfeld@kdis.ac.kr}{\nolinkurl{merfeld@kdis.ac.kr}}}}
\date{2023-04-12}

\begin{document}
\maketitle
\begin{abstract}
\noindent I document externalities in the agricultural sector of India due to air pollution. Using variation in pollution induced by changes in wind, I show that higher levels of particulate matter lead to decreased agricultural productivity, with changes in productivity of 10 percent being common. Combining wind direction with the rollout of coal plants, results indicate pollution from coal plants has a larger effect on agricultural productivity than other types of pollution. Given that the agricultural sector is a refuge for the poor in many developing countries, these results suggest that the effects of pollution may hit the poorest particularly hard.\\
\strut \\
\textbf{\textit{Keywords}}: pollution, productivity, agriculture, labor, India\\
\textbf{\textit{JEL Codes}}: H40, I15, J22, O13, Q52, Q53
\end{abstract}
